\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-03-11 forwards-and-backwards.tex
stationary dist corollary

\begin{corollary} \label{cor:lineages_simple}
    In addition to the assumptions of Theorem~\ref{thm:lineages},
    if the dispersal process is isotropic
    (i.e., $\covq = \sigma^2 I$ and $\meanq$ is constant),
    then
    \begin{equation}
        \Lgen_\varphi f
        =
        r \gamma
        \left(
            \sigma^2 \Delta f
            +
            \left(
                2 \sigma^2 \grad \log(\gamma \varphi)
                - \meanq
            \right)
            \cdot \grad f
        \right) .
    \end{equation}
\end{corollary}

In other words, 
the lineage behaves as a diffusion driven by Brownian motion 
run at speed $\sigma^2$ multiplied by the local per-capita production of 
mature offspring 
($r \gamma$)
in a potential equal to the local output of juveniles tilted by migration bias
($\varphi_s \gamma \exp(-\meanq \cdot x / 2 \sigma^2)$). In particular,
lineages are drawn to regions of high fecundity (production of juveniles), but their
speed is determined by the rate of production of mature offspring.
This can be compared to Remark~\ref{ancestral lineages: first guess}.

\begin{corollary}
    \comment{TODO: write corollary saying what's in the previous sentence 
    and what the stationary distribution is?}
\end{corollary}

@@

>>>>

\begin{corollary} \label{cor:lineages_simple}
    In addition to the assumptions of Theorem~\ref{thm:lineages},
    if the covariance of the dispersal process is isotropic
    (i.e., $\covq = \sigma^2 I$),
    then
    \begin{equation}
        \Lgen_\varphi f
        =
        r \gamma
        \left(
            \sigma^2 \Delta f
            +
            \left(
                2 \sigma^2 \grad \log(\gamma \varphi)
                - \meanq
            \right)
            \cdot \grad f
        \right) .
    \end{equation}
    (However, $\meanq$ can still depend on location.)
\end{corollary}

In other words, 
the lineage behaves as a diffusion driven by Brownian motion 
run at speed $\sigma^2$ multiplied by the local per-capita production of 
mature offspring ($r \gamma$)
in a potential tilted by migration bias
($(\varphi_s \gamma)^2 \exp(-\meanq \cdot x / \sigma^2)$).
In particular, lineages are drawn to regions of high fecundity (production of juveniles), but their
speed is determined by the rate of production of mature offspring.
This can be compared to Remark~\ref{ancestral lineages: first guess}.

\begin{corollary} \label{cor:stationary_dist}
    In addition to the assumptions of Corollary~\ref{cor:lineages_simple},
    the population process is stationary (so $\varphi_t \equiv \varphi$),
    and $\meanq(x) = \grad h(x)$ for some function $h$,
    then $Y$ is reversible with respect to
    \begin{align}
        \pi(x)
        =
        \frac{\gamma}{r} \varphi(x)^2 e^{-h(x) / \sigma^2} .
    \end{align}
\end{corollary}

\comment{I'm a bit surprised by the factor of 2 (that manifests as $\varphi^2$),
    but I've checked where it comes from pretty carefully, and don't see anything wrong.}
Long-term fitness of an individual
is proportional to the fraction of lineages from the distant future that pass through the individual,
and hence the total long-term fitness at a location is proportional to
the stationary distribution of $Y$ there, if it exists.
Therefore,
if $\pi$ is integrable then the per-capita long-term fitness of an individual at $x$
is proportional to $\pi(x) / \varphi(x)$.

@@

\begin{proof}[of Corollary~\ref{cor:stationary_dist}]
For the moment, we will write $r(x)$ for $r(x, \eta)$
and similarly for $\gamma(x)$.
First, note that in this case
$$
    \Lgen f = \sigma^2 r \gamma \left(
        \Delta f
        + \grad ( 2 \log(\gamma \varphi) - h ) \cdot \grad f
    \right) ,
$$
and that
$$
    \int_{\IR^d} e^{H(x)} f(x) (\Delta + \grad H(x) \cdot \grad) g(x) dx
    =
    - \int_{\IR^d} e^{H(x)} \left\{
        \grad f(x) \cdot \grad g(x)
    \right\} dx .
$$
Therefore, by choosing $H(x) = 2 \log(\gamma(x) \varphi(x)) - h(x)$
and
$$ \pi(x)
    = e^{H(x)} / (\sigma^2 r(x) \gamma(x))
    = \frac{\gamma(x) \varphi(x)^2 e^{-h(x)}}{\sigma^2 r(x)} ,
$$
we have that
$$
    \int_{\IR^d} \pi(x) f(x) \Lgen g(x) dx
    =
    - \int_{\IR^d} e^{H(x)} \grad f(x) \cdot \grad g(x) dx .
$$
Since this is symmetric, the process $Y$ is reversible with respect to $\pi$;
the constant factor of $\sigma^2$ does not affect the result.
\end{proof}
